\subsection{}

Prove the Subject Reduction theorem by induction on the height of the
derivation of the judgement $\Gamma \vdash M: T$

The Subject Reduction theorem states that if $\Gamma \vdash M: T$ is derivable and $M \to M'$
then $\Gamma \vdash M': T$ is derivable.\\
The proof will proceed by induction on the height of the derivation tree of the judgement
$\Gamma \vdash M: T$

\paragraph*{Base cases: $h = 1$:\\}

The base cases are those where the last typing rule is an axiom:
\begin{itemize}
	\item (T-TRUE): $true \not\to$ thus vacuously true
	\item (T-FALSE): $false \not\to$ thus vacuously true
	\item (T-INT): $n \not\to$ thus vacuously true
	\item (T-VAR): $x \not\to$ thus vacuously true
\end{itemize}

\paragraph*{Inductive cases: $h = h_0 + 1$:\\}

The inductive hypothesis is that for any derivable judgement $\Gamma \vdash M: T$ of height
$h \le h_0$ if $M \to M'$ then $\Gamma \vdash M': T$ is derivable.\\
The inductive cases will be distinguished based on the last typing rule applied to the
derivation tree:

\begin{itemize}
	\item (T-SUM): $\Gamma \vdash M: Nat$ derivable with $M = \SUM{M_1}{M_2}$;
	      by Inversion Lemma:
	      \begin{itemize}
		      \item $\Gamma \vdash M_1: Nat$ derivable
		      \item $\Gamma \vdash M_2: Nat$ derivable
	      \end{itemize}
	      if $M \to M'$ to get $M'$ it was applied one of the following operational semantics rules:
	      \begin{itemize}
		      \item (SUM): $M = \SUM{n_1}{n_2} \to n = M'$, thus by (T-INT) rule
		            $\Gamma \vdash n: Nat$ derivable
		      \item (SUM-LEFT): $M \to \SUM{M_1'}{M_2} = M'$,
		            since $\Gamma \vdash M_1: Nat$ derivable
		            by Inv. Lemma with height $h_0$ the inductive hypothesis can be applied, hence
		            $\Gamma \vdash M_1': Nat$ derivable. Thus by (T-SUM) rule $\Gamma \vdash M': Nat$
		            derivable
		      \item (SUM-RIGHT): $M \to \SUM{v}{M_2'}$, since $\Gamma \vdash M_2: Nat$ derivable
		            by Inv. Lemma with height $h_0$ the inductive hypothesis can be applied, hence
		            $\Gamma \vdash M_2': Nat$ derivable. Thus by (T-SUM) rule $\Gamma \vdash M': Nat$
		            derivable
	      \end{itemize}
	\item (T-IFTHENELSE): $\Gamma \vdash M: T$ derivable with $M = \IF{M_1}{M_2}{M_3}$;
	      by Inversion Lemma:
	      \begin{itemize}
		      \item $\Gamma \vdash M_1: Bool$ derivable
		      \item $\Gamma \vdash M_2: T$ derivable
		      \item $\Gamma \vdash M_3: T$ derivable
	      \end{itemize}
	      if $M \to M'$ to get $M'$ it was applied one of the following operational semantics rules:
	      \begin{itemize}
		      \item (IF-TRUE): $\IF{true}{M_2}{M_3} \to M_2 = M'$ with $\Gamma \vdash M': T$
		            derivable by Inversion Lemma
		      \item (IF-FALSE): $\IF{false}{M_2}{M_3} \to M_3 = M'$ with $\Gamma \vdash M': T$
		            derivable by Inversion Lemma
		      \item (IF): $M \to \IF{M_1'}{M_2}{M_3} = M'$ since $\Gamma \vdash M_1: Bool$ derivable
		            by Inv. Lemma with height $h_0$ the inductive hypothesis can be applied, hence
		            $\Gamma \vdash M_1': Bool$ derivable. Thus by (T-IFTHENELSE) rule
		            $\Gamma \vdash M': T$ derivable
	      \end{itemize}
	\item (T-FUN): $M = \FN{x: T_1}{M_1} \not\to$ thus vacuously true
	\item (T-APP): $\Gamma \vdash M: T$ derivable with $M = \APP{M_1}{M_2}$;
	      by Inversion Lemma $\exists\,T_1$ such that:
	      \begin{itemize}
		      \item $\Gamma \vdash M_1: T_1 \to T$ derivable
		      \item $\Gamma \vdash M_2: T_1$ derivable
	      \end{itemize}
	      if $M \to M'$ to get $M'$ it was applied one of the following operational semantics rules:
	      \begin{itemize}
		      \item (BETA): $M = \APP{(\FN{x: T_1}{M_{1_1}})}{v} \to M\SUBST{x}{v}$, by Inversion
		            Lemma\\
		            $\Gamma \vdash \FN{x:T_1}{M_{1_1}}: T_1 \to T$ and $\Gamma \vdash v: T_1$.
		            By Inversion Lemma once again on $M_1$ $\Gamma, x: T_1 \vdash M_{1_1}: T$.
		            Now by Substitution Lemma $\Gamma \vdash M\SUBST{x}{v}\;(= M'): T$
		      \item (APP1): $M \to \APP{M_1'}{M_2} = M'$ since $\Gamma \vdash M_1: T_1 \to T$
		            derivable by Inv. Lemma with height $h_0$ the inductive hypothesis can be applied,
		            hence $\Gamma \vdash M_1': T_1 \to T$ derivable. Thus by (T-APP) rule
		            $\Gamma \vdash M': T$ derivable
		      \item (APP2) $M \to \APP{v}{M_2'} = M'$ since $\Gamma \vdash M_2: T_1$ derivable by
		            Inv. Lemma with height $h_0$ the inductive hypothesis can be applied,
		            hence $\Gamma \vdash M_2': T_1$ derivable. Thus by (T-APP) rule
		            $\Gamma \vdash M': T$ derivable
	      \end{itemize}
\end{itemize}
