\subsection{}

\begin{enumerate}
	\item Add to the language $\mathcal{L}$ a subtraction operation.
	      What needs to be changed, the syntax of terms? The semantics?
	      The syntax of values? The notion of free-variables and substitution?
	\item Define the reduction relation so that integer operations occur with a
	      right-to-left strategy.
	\item The given reduction relation defines an efficient strategy for the
	      evaluation of the if-then-else term, i.e. allowing only the branch chosen
	      by the evaluation of the Boolean guard to be evaluated. Redefine the
	      operational semantics of the language so that it adopts an inefficient
	      strategy for the if-then-else construct, evaluating both branches of the
	      conditional construct before selecting one of them. Is the relationship
	      thus defined ambiguous (nondeterministic)?
	      What is the semantics of the term $\IF{false}{\SUM{3}{true}}{7}$?
\end{enumerate}

\begin{enumerate}
	\item In order to add the subtraction operation to $\mathcal{L}$ the syntax
	      of the terms must be extended by adding the term $M_1 - M_2$, the
	      semantics must be extended as follows
	      \[
		      \infer[n = n_1 -_\mathbb{Z} n_2]{n_1 - n_2 \to n}{}
	      \]
	      \[
		      \infer[\text{\small{(SUB-L)}}]{M_1 - M_2 \to M_1' - M_2}{M_1 \to M_1'}
		      \qquad
		      \infer[\text{\small{(SUB-R)}}]{v - M \to v - M'}{M \to M'}
	      \]
	      The syntax of values remains unchanged since $n$ are integer constants,
	      while the notions of free-variables and substitution must be enlarged
	      with the newly added term as follows
	      \begin{align*}
		       & \fv{M_1 - M_2} = \fv{M_1} \cup \fv{M_2}                     \\
		       & (M_1 - M_2)\SUBST{x}{N} = M_1\SUBST{x}{N} - M_2\SUBST{x}{N}
	      \end{align*}
	\item The SUM-LEFT and SUM-RIGHT rules will be replaced respectively by the rules\\
	      SUM-LEFT' and SUM-RIGHT' defined as follows:\\
	      \[
		      \infer[\text{\small{SUM-RIGHT'}}]{\SUM{M_1}{M_2} \to \SUM{M_1}{M_2'}}{M_2 \to M_2'}
		      \qquad
		      \infer[\text{\small{SUM-LEFT'}}]{\SUM{M}{v} \to \SUM{M'}{v}}{M \to M'}
	      \]
	\item The IF-TRUE and IF-FALSE axioms must be replaced by the followings:\\
	      \[
		      \infer[\text{\small{IF-TRUE'}}]{\IF{true}{v_1}{v_2} \to v_1}{}
		      \qquad
		      \infer[\text{\small{IF-FALSE'}}]{\IF{false}{v_1}{v_2} \to v_2}{}
	      \]
	      And the following two rules must be added:\\
	      \[
		      \infer[\text{IF-L}]{\IF{v}{M_1}{M_2} \to \IF{v}{M_1'}{M_2}}{M_1 \to M_1'}
	      \]
	      \[
		      \infer[\text{IF-R}]{\IF{v}{v_1}{M} \to \IF{v}{v_1}{M'}}{M \to M'}
	      \]
	      The relationship as defined would still be deterministic and the term
	      $\IF{false}{\SUM{3}{true}}{7}$ would be stuck since the left branch
	      isn't a value and can't be further reduced by the given operational semantics.
\end{enumerate}
